Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and music aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-04-10 introduction/Subsections/Motivation.tex
Updates from Overleaf

I have consistently been inspired by musicians and researchers who strive to uncover musical ground truth within the existing tonal paradigm, such as Heinrich Schenker \cite{}, or by challenging it, as exemplified by George Russell \cite{LydianRussell}. Both approaches aim to identify abstract concepts that reinforce or disrupt the tonal foundation. Their ultimate goal is to advance the tonal landscape, providing musicians with a dependable playground for growth, development, and further understanding of tonal paradigms.

@@

Anstieg (Initial Ascent): Ascending motion leading to the primary tone of the fundamental line.
Arpeggiation (Brechung): Elementary elaboration of a harmony.
Background (Hintergrund): The structural level of the fundamental structure.
Bass Arpeggiation (Bassbrechung): Bass pattern I-V-I forms the harmonic content of the background of tonal musical pieces.
Chord of Nature (Klang): The complex sound consisting of the first five harmonic series notes suggests a model for the major triad.
Diminution: The process by which an interval formed by notes of longer value is expressed in notes of smaller value.
Foreground (Vordergrund): The surface level of musical structure in Schenkerian analysis.
Fundamental Line (Urlinie): The melodic aspect of the fundamental structure, a stepwise descent from one of the triad notes to the tonic.
Middleground (Mittelgrund): An intermediate structural level between the background and foreground in Schenkerian analysis.
Prolongation (Auskomponierung): The process in tonal music through which a pitch, interval, or consonant triad can govern music spans when not physically sounding.
Structural Level (Schicht): Refers to the different levels of musical structure in Schenkerian analysis, including background, middle-ground, and foreground.
Tonal Space: A general principle of Schenkerian analysis where intervals between the notes of the tonic triad form a tonal space filled with passing and neighbor notes.
Unfolding (Ausfaltung): The transformation of a single chord into a horizontal succession.
Voice Leading: The study of the principles that govern the progression of the component voices of the composition both separately and in combination.
The glossary also includes related articles, references, and external links for further reading and understanding Schenkerian analysis.

George Russell's Lydian Chromatic Concept:
George Russell's Lydian Chromatic Concept of Tonal Organization is an innovative approach to understanding and organizing musical relationships. Introduced in the 1950s, the theory asserts that the Lydian mode, rather than the traditional major scale, is the accurate parent scale in Western music. The Lydian scale is characterized by a raised fourth degree, giving it a brighter and more consonant sound. Russell's concept revolves around "chord modes," derived from a chord's parent Lydian scale. This system allows for greater harmonic flexibility and a broader range of tonal colors, influencing the development of jazz and contemporary music.

>>>>

I have consistently been inspired by musicians and researchers who strive to uncover musical ground truth within the existing tonal paradigm, such as Heinrich Schenker \cite{}, or by challenging it, as exemplified by George Russell \cite{LydianRussell}. Both approaches aim to identify abstract concepts that reinforce or disrupt the tonal foundation. Their ultimate goal is to advance the tonal landscape, providing musicians with a dependable playground for growth, development, and further understanding tonal paradigms.

@@

Anstieg (Initial Ascent): Ascending motion leading to the primary tone of the fundamental line.
Arpeggiation (Brechung): Elementary elaboration of a harmony.
Background (Hintergrund): The structural level of the fundamental structure.
Bass Arpeggiation (Bassbrechung): Bass pattern I-V-I forms the harmonic content of the background of tonal musical pieces.
Chord of Nature (Klang): The complex sound of the first five harmonic series notes suggests a model for the major triad.
Diminution: The process by which an interval formed by notes of longer value is expressed in notes of smaller value.
Foreground (Vordergrund): The surface level of musical structure in Schenkerian analysis.
Fundamental Line (Urlinie): The melodic aspect of the fundamental structure, a stepwise descent from one of the triad notes to the tonic.
Middleground (Mittelgrund): An intermediate structural level between the background and foreground in Schenkerian analysis.
Prolongation (Auskomponierung): The process in tonal music through which a pitch, interval, or consonant triad can govern music spans when not physically sounding.
Structural Level (Schicht): Refers to the different levels of musical structure in Schenkerian analysis, including background, middle-ground, and foreground.
Tonal Space: A general principle of Schenkerian analysis where intervals between the notes of the tonic triad form a tonal space filled with passing and neighbor notes.
Unfolding (Ausfaltung): The transformation of a single chord into a horizontal succession.
Voice Leading: The study of the principles that govern the progression of the component voices of the composition both separately and in combination.
The glossary also includes related articles, references, and external links for further reading and understanding Schenkerian analysis.

%%%%%%%%%%%%%
%MESOSTRUC
%%%%%%%%%%%%%

Mesostructures \cite{Mesostructures2023} are intermediate levels of articulation between waveshapes' microstructure and musical forms' macrostructure. Examples of mesostructures include melody, arpeggios, syncopation, polyphonic grouping, and textural contrast. The article notes that mesostructures have received limited attention in deep learning, as current models only focus on microstructures. The article proposes a new approach to mesostructural audio modeling, which involves a differentiable arpeggiator and time-frequency scattering. The article also discusses the need for a time-invariant and multiscale differentiable time-frequency model of similarity at the level of both local spectra and spectrotemporal modulations.

### 2023-04-10 methods/subsections/machine learning.tex
Updates from Overleaf

\textbf{1D Convolution Example:} Given a 1D input signal $I = [i_1, i_2, ..., i_n]$ and a filter (or kernel) $K = [k_1, k_2, ..., k_m]$, the convolution operation is defined as $(I * K)j = \sum{u=1}^{m} I_{j-u+1} K_u$, where $j$ ranges from $1$ to $n-m+1$, and $I * K$ represents the resulting feature vector after applying the convolution. The objective of the 1D convolution operation is to slide the filter over the input signal and compute the element-wise product and sum for each local region, effectively learning and detecting local patterns in the input data.

>>>>

\textbf{Convolution:} It involves sliding a small window, called a filter or kernel, over input data to compute the dot product between the kernel and each local region of the data. The result is a new feature map highlighting patterns and essential information.

Given a 2D input data represented as a matrix $I = [i_{ij}]{m \times n}$ and a filter $K = [k{ij}]_{p \times q}$, the convolution operation is defined as:

$$(I * K)_{kl} = \sum_{i=1}^{p} \sum_{j=1}^{q} i_{k+i-1, l+j-1}k_{ij}$$

Here, $k$ ranges from $1$ to $m-p+1$ and $l$ ranges from $1$ to $n-q+1$. The resulting feature map after applying the convolution is denoted by $(I * K)_{kl}$.

The convolution operation slides the kernel $K$ over the input image $I$, performing an element-wise multiplication and sum of the local region centered at pixel $(k,l)$ to compute the value of each element in the output feature map, $(I * K)_{kl}$. This process is repeated for each pixel $(k,l)$, resulting in a new feature map that captures local patterns in the input image.

\input{figures/neural networks/convolution}

### 2023-04-10 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

\subsection{Contrastive Self-Supervised Learning for Musical Representations (CSSLMR)}

Contrastive Self-Supervised Learning for Musical Representations (CSSLMR) is an approach that combines the principles of Self-Supervised Learning (SSL) and Contrastive Learning of Musical Representations (CLMR) to learn discriminative and valuable representations of music without relying on explicit labels. By leveraging the structure and content of music and contrasting different augmentations or transformations of the same musical piece against randomly (or not) selected pieces, CSSLMR models can effectively capture the underlying structure and known invariance in the data.

>>>>

\subsection{Contrastive Self-Supervised Learning of Musical Representations (CSSLMR)}

Contrastive Self-Supervised Learning of Musical Representations (CSSLMR) is an approach that combines the principles of Self-Supervised Learning (SSL) and Contrastive Learning of Musical Representations (CLMR) to learn discriminative and valuable representations of music without relying on explicit labels. By leveraging the structure and content of music and contrasting different augmentations or transformations of the same musical piece against randomly (or not) selected pieces, CSSLMR models can effectively capture the underlying structure and known invariance in the data.
